---
title: "Personnal Glucose Readings Analysis"
author: "Craig W. Slinkman"
date: "`r Sys.Date()`"
output: html_document
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction  

## Initialization

```{r  Initialization}
require(here)                          # Needed to determine project location.
require(tidyverse)                     # I live in the tidyverse.
require(forcats)                       # Used to handle categorical variables
require(lubridate)                     # Needed for date manipulation
require(flextable)                     # Provides table display functionality.
require(cowplot)                       # For my favorite plot look.

# Define require functions.
source("D:/R-Projects/Rsource/plot_distribution.R")              # Relative ...
                                                                 #   frequency 
                                                                 #   dist.
source("D:/R-Projects/Rsource/plot_qq_norm.R")                   # Draw qq plot.
source("D:/R-Projects/glucose/functions/compute_mode.R")         # Compute mode.
source("D:/R-Projects/glucose/functions/compute_iqr.R")         # Compute iqr. 
source("D:/R-Projects/glucose/functions/compute_statistics.R")  # Compute
                                                                #  summary
                                                                #  statistics.
source("D:/R-Projects/glucose/functions/plot_time_series.R")    # Plot time-
                                                                #  series.

theme_set(theme_cowplot())             # Set the cowplot_theme as the default.
```

## Locate data  directory and read files 

```{r Determine_file_path_to_the_data}
project <- here()                      # Determine location in file system.

file <- 
  file.path( project,                  # Construct file path to data file.               
             "data",
             "diabetes.csv" )

Readings <-                           # Read file into tibble. 
  read_csv( file )

spec( Readings )                      # Show tibble column specifications.


################################################################################
# Change date_time from character data to a date_time object.
################################################################################

Readings <- 
  Readings %>%
    mutate( date_time = 
              parse_date_time( 
                date_time,
                "YmdHM",
                locale = Sys.getlocale("LC_TIME" )))

################################################################################
# We only use the last 180 days measurements.
################################################################################

Readings <-
  Readings %>% 
    filter( date_time > now() - 180 * 24 * 60 * 60 )


```  

## Last 14 (One week) observarions

Note that the glucose readings below showing the readings in reverse chronological order.  Therefore, the last observation is shown first.

```{r Display_last_14_observations}
Readings %>% 
  slice_tail( n = 14 ) %>% 
  arrange( desc( date_time )) %>% 
    flextable() %>% 
      autofit()
```

  
## Display barchart of time descriptions

We draw a barchart of the frequency of observations taken in the morning(AM)
and the number of observarions taken in the evening(PM)

```{r Draw_bar_chart_of_activities}
this_title <- 
  paste( "Activities as of ", now() )

Readings <-
  Readings %>% 
    mutate( meridian = 
            if_else( hour(date_time) < 12,
                        "AM","PM"))


ggplot( Readings,
        aes( x = meridian )) +
  geom_bar( color = "black", 
           fill  = "green" ) +
    xlab( "When measured" ) +
    ylab( "Frequency" ) +
    ggtitle( this_title )

```

## Time Series plot of glucose readings

```{r plot_last_180_days_observations}
y_scale <- seq(from = 50,                   # glucose glucose marks and labels.
                to  = 200,
                by = 10 )

this_title <- 
  paste( "Blood Glucose readings as of"     # Create plot title.
         , now())

ggplot( Readings,                           # Display plot.
        aes( x   = date_time,
             y   = glucose )) +
  geom_line() +
  xlab( "Date-Time" ) +
  ylab( "Glucose(mg/dl)") +
  ggtitle( this_title ) +
  scale_y_continuous( breaks = y_scale, 
                      labels = y_scale,
                      limits = c( 60, 200 ))

```  

## Boxplot of glucose readings by time description

```{r Darw_boxplot_of_glucose_by_time}
this_title <-
  paste( "Blood Glucose readings by when in day taken\n as of",
         now() )

ggplot( Readings,
        aes( x = meridian,
             y = glucose )) +
        geom_boxplot( na.rm          = TRUE,
                      outlier.colour = "red", 
                      outlier.shape  = 16,
                      outlier.size   = 2, 
                      notch          =TRUE ) +
  scale_y_continuous( breaks = y_scale, 
                      labels = y_scale,
                      limits = c( 60, 200 )) +
  ggtitle( this_title )
```
  
  
## Extract gluocse vector from the readings tibble

```{r Extract_glucose}
glucose <-                                  # Extract glucose
  Readings %>% 
    pull(glucose)

glucose <-                                  # Remove missing values.
  glucose[!is.na(glucose)]

n <- length(glucose)                        # Determine length of glucose
                                            # vector
```

## Rleative frequency distribution with density smoother

```{r Plot_histogram_of_glucose}
this_title <- 
  paste( "Relative frquency distribution with smoothed density",
         "\nas of", now())
plot_distribution( glucose, 
                   "Glucose(mg/dl)", 
                   this_title )
```

## Normal Quantile-Quentile plot (qq-plot)

```{r plot_qq_norm_glucose}
plot_qq_norm( glucose, R = 100,
              variable_name = "Glocose(mg/dl)",
              "QQ Plot of glucose")
```
  
## Time series plot with lowess smoothing function
  
```{r Draw loess_plot_of_glucose}
TS <-                                       # Extract time series
  Readings %>% 
      mutate( date = date(date_time))
      
  
n <- dim(TS)[1]                             # Create time series time index.
index <- 1:n

TS <-
  TS %>% 
    add_column( index  )

s <-
  loess( glucose ~ index,
         data = TS,
         span = 2/3 )

smooth <- predict(s)

TS <-
  TS %>% 
    add_column( smooth  )
                                 # Display tibble.

ggplot( TS,                      # Display time series plot.
        aes( x   = date_time,
             y   = glucose )) +
  geom_line() +
  geom_line( aes( x =  date_time, 
                  y = smooth )) +
  xlab( "Date-Time") +
  ylab( "Smoothed glucose readings(mg/dL)") +
  ggtitle( "Smoothed Glucose readings by Date")
 
```

## Descriptive statistics

### Glucose percentiles

```{r Compute_180_day_percentile}
 Probability <-                                
  c( 0.005, 
     0.010, 
     0.025, 
     0.050,
     0.250,
     0.500,
     0.750,
     0.950,
     0.975,
     0.990,
     0.995 )

Values <- quantile( glucose, 
               probs = Probability,
               omit.na = TRUE )

Percentiles <-
  tibble( Probability,
          Values )


Percentiles %>% 
  flextable() %>% 
    add_header(
      values =  list( 
       Probability  = "Glucose Percentiles" )) %>% 
       merge_at( i = 1, j = 1:2, part = "header") %>% 
       align( i = 1,
              j = 1,
       align = "left",
       part = "header" ) %>% 
    autofit() %>% 
    theme_box()
```


### Compute and display descriptive statistics by days

```{r display_summary_statistics by days}
Statistics007 <- compute_statistics( Readings, 7  ) 
Statistics014 <- compute_statistics( Readings, 14 )
Statistics030 <- compute_statistics( Readings, 30 )
Statistics090 <-compute_statistics( Readings, 90 )
Statistics180 <- compute_statistics( Readings, 180 )
  
Report <-                                   # Create report structure
  bind_rows( Statistics007,
             Statistics014,
             Statistics030,
             Statistics090,
             Statistics180 )

Report %>%                                 # Display report as a flextable.
  flextable() %>% 
  autofit()
```

## Analysis by number of days

### Last 7 day's glucose readings

```{r plot_last_007_days_readings}
plot_time_series( Readings, 7 )            # Plot last seven days readings.
Statistics007 %>%                          # Display report as a flextable.
  flextable() %>% 
  autofit()
```

### Last 14 day's glucose readings 

```{r}
plot_time_series( Readings, 14 )            # Plot last seven days readings.

Statistics014 %>% 
  flextable() %>% 
  autofit()
```
   
### Last 30 day's glucose readings 

```{r display_lst_30_days_observarion}
plot_time_series( Readings, 90 )            # Plot last seven days readings.

Statistics030 %>% 
  flextable() %>% 
  autofit()
```

### Last 90 day's glucose readings 

```{r display_lst_90_days_observarion}
plot_time_series( Readings, 90 )            # Plot last seven days readings.

Statistics090 %>% 
  flextable() %>% 
  autofit()
```