---
title: "Personnal Glucose Readings Analysis"
author: "Craig W. Slinkman"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction  

## Initialization

```{r}
require(here)                          # Needed to determine project location.
require(tidyverse)                     # I live in the tidyverse.
require(forcats)                       # Used to handle categorical variables
require(lubridate)                     # Needed for date manipulation
require(flextable)                     # Provides table display functionality.
require(cowplot)                       # For my favorite plot look.

# Define require functions.
source("D:/R-Projects/Rsource/plot_distribution.R")              # Relative 
                                                                 #   frequency 
                                                                 #   dist.
source("D:/R-Projects/Rsource/plot_qq_norm.R")                   # Normal
                                                                 #  plot.
source("D:/R-Projects/RStatistics/compute_mode.R")               # Compute mode.
source("D:/R-Projects/Diabetes/functions/compute_iqr.R")         # Compute mode. 
source("D:/R-Projects/Diabetes/functions/compute_statistics.R")  # For stats.




theme_set(theme_cowplot())             # Set the cowplot_theme as the default.
```

## Locate data  direcotry and read files 

```{r}
project <- here()                      # Determine location in file system.

file <- 
  file.path( project,                  # Construct file path to data file.               
             "data",
             "diabetes.csv" )

Readings <-                           # Read file into tibble. 
  read_csv( file )

spec( Readings )                      # Show tibble column specifications.


################################################################################
# Change date_time from character data to a date_time object.
################################################################################

Readings <- 
  Readings %>%
    mutate( date_time = 
              parse_date_time( 
                date_time,
                "YmdHM",
                locale = Sys.getlocale("LC_TIME" )))

################################################################################
# Change the names of the variables to make our data list and graphics names
# start with capital letters.  We only use the last 180 days measurements.
################################################################################

Readings <-
  Readings %>% 
    rename( "Date_time" = "date_time",
            "Description" = "description",
            "Glucose"     = "glucose",
            "Finger"      = "finger" ) %>% 
    filter( Date_time > now() - 180 * 24 * 60 * 60 )


```  

## Last 10 observarions

```{r}
Readings %>% 
  slice_tail( n = 10 ) %>% 
    flextable() %>% 
      autofit()
```

  
## Display barchart of time descriptions

```{r}
dl <- 
  c( "Morning", "Post Ride", "Evening" )    # Define descriptive order

Readings <-
  Readings %>% 
   mutate( Description =                    # Reorder factors by time.
            fct_relevel( Description, 
                         dl ))

this_title <- 
  paste( "Activities as of ", now() )

ggplot( Readings,
        aes( x = Description )) +
  geom_bar( color = "black", 
           fill  = "green" ) +
    xlab( "Activity DeScription" ) +
    ylab( "Frequency" ) +
    ggtitle( this_title )

```

## Time Series plot of glucose readings

```{r}
y_scale <- seq(from = 50,                   # glucose glucose marks and labels.
                to  = 200,
                by = 10 )

this_title <- 
  paste( "Blood Glucose readings as of"     # Create plot title.
         , now())

ggplot( Readings,                           # Display plot.
        aes( x   = Date_time,
             y   = Glucose )) +
  geom_line() +
  # geom_point( aes( color = Description),
  #             size = 2 ) +
  xlab( "Date-Time" ) +
  ylab( "Glucose(mg/dl)") +
  ggtitle( this_title ) +
  scale_y_continuous( breaks = y_scale, 
                      labels = y_scale,
                      limits = c( 60, 200 ))

```  

## Boxplot of glucose readings by time description

```{r}
this_title <-
  paste( "Blood Clucose readings by time description\n as of",
         now() )

ggplot( Readings,
        aes( x = 
               fct_relevel( Description, dl ),
             y = Glucose )) +
        geom_boxplot( na.rm          = TRUE,
                      outlier.colour = "red", 
                      outlier.shape  = 16,
                      outlier.size   = 2, 
                      notch          =TRUE ) +
  scale_y_continuous( breaks = y_scale, 
                      labels = y_scale,
                      limits = c( 60, 200 )) +
  ggtitle( this_title )
```
  
  
## Extract gluocse vector from the readings tibble

```{r}
Glucose <-                                  # Extract glucose
  Readings %>% 
    pull(Glucose)

Glucose <-                                  # Remove missing values
  Glucose[!is.na(Glucose)]

n <- length(Glucose)                        # Determine length of glucose
                                            # vector
```

## Rleative frequency distribution with density smoother

```{r}
this_title <- 
  paste( "Relative frquency distribution with smoothed density",
         "\nas of", now())
plot_distribution( Glucose, 
                   "Glucose(mg/dl)", 
                   this_title )
```

## Normal Quantile-Quentile plot (qq-plot)

```{r}
plot_qq_norm( Glucose, R = 100,
              variable_name = "Glocose(mg/dl)",
              "QQ Plot of glucose")
```
  
## Time series plot with lowess smoothing function
  
```{r}
TS <-                                       # Extract time series
  Readings %>% 
    filter( Description == "Morning" |
            Description == "Evening" ) %>% 
      mutate( date = date(Date_time))
      
  
n <- dim(TS)[1]                             # Create time series time index.
Index <- 1:n

TS <-
  TS %>% 
    add_column( Index  )

s <-
  loess( Glucose ~ Index,
         data = TS,
         span = 2/3 )

Smooth <- predict(s)

TS <-
  TS %>% 
    add_column( Smooth  )
                                 # Display tibble.

ggplot( TS,                      # Display time series plot.
        aes( x   = Date_time,
             y   = Glucose )) +
  geom_line() +
  geom_line( aes( x =  Date_time, 
                  y = Smooth )) +
  xlab( "Date Time") +
  ggtitle( "Smoothed Glucose readings by Date")
 
```

## Descriptive statistics

### Glucose percentiles



```{r}
Probability <-
  c( 0.005, 0.010, 0.025, 0.050,
        seq( from = 0.10, 
             to   = 0.90, 
             by = 0.10 ),
    0.950, 0.975, 0.995 )

Percentile <- quantile( Glucose, 
               probs = Probability,
               omit.na = TRUE )

percent <- names(q)

Percentiles <-
  tibble( Probability,
          Percentile )

Percentiles %>% 
  flextable() %>% 
    autofit() %>% 
    theme_box()
```


### Compute and display descriptive statistics by days

```{r}

Statistics007 <- compute_statistics( Readings, 7  )
Statistics014 <- compute_statistics( Readings, 14 )
Statistics030 <- compute_statistics( Readings, 30 )
Statistics090 <-compute_statistics( Readings, 90 )
Statistics180 <- compute_statistics( Readings, 180 )
  
Report <- 
  bind_rows( Statistics007,
             Statistics014,
             Statistics030,
             Statistics090,
             Statistics180 )

header_values <-                            # Create header for flextable.
  names( Report )

header_values[11] <-
  "Precent\nhigh" 


Report %>% 
  flextable() %>% 
    add_header_row( top = TRUE,
                    values = header_values ) %>% 
  autofit()

```

## Apendencix: The data set

```{r}
Readings %>% 
  flextable %>% 
    autofit()
```


